---
title: "Exploratory data analysis for Rossmann Store Sales"
author: "Juan L. Kehoe"
date: "10/12/2015"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic information 
Competition link: [https://www.kaggle.com/c/rossmann-store-sales](https://www.kaggle.com/c/rossmann-store-sales)    
Data download link: [https://www.kaggle.com/thie1e/exploratory-analysis-rossmann/data](https://www.kaggle.com/thie1e/exploratory-analysis-rossmann/data)   

## Purpose of the project
Predict 6 weeks of daily sales for 1,115 stores located across Germany.

## Data provided:
You are provided with historical sales data for 1,115 Rossmann stores. The task is to forecast the "Sales" column for the test set. Note that some stores in the dataset were temporarily closed for refurbishment.    

### Files
- train.csv - historical data including Sales  
- test.csv - historical data excluding Sales   
- store.csv - supplemental information about the stores  

### Data fields
Most of the fields are self-explanatory. The following are descriptions for those that aren't.  

- Id - an Id that represents a (Store, Date) duple within the test set  
- Store - a unique Id for each store  
- Sales - the turnover for any given day (this is what you are predicting)  
- Customers - the number of customers on a given day  
- Open - an indicator for whether the store was open: 0 = closed, 1 = open  
- StateHoliday - indicates a state holiday. Normally all stores, with few exceptions, are closed on state holidays. Note that all schools are closed on public holidays and weekends. a = public holiday, b = Easter holiday, c = Christmas, 0 = None  
- SchoolHoliday - indicates if the (Store, Date) was affected by the closure of public schools  
- StoreType - differentiates between 4 different store models: a, b, c, d  
- Assortment - describes an assortment level: a = basic, b = extra, c = extended  
- CompetitionDistance - distance in meters to the nearest competitor store  
- CompetitionOpenSince[Month/Year] - gives the approximate year and month of the time the nearest competitor was opened  
- Promo - indicates whether a store is running a promo on that day  
- Promo2 - Promo2 is a continuing and consecutive promotion for some stores: 0 = store is not participating, 1 = store is participating  
- Promo2Since[Year/Week] - describes the year and calendar week when the store started participating in Promo2
- PromoInterval - describes the consecutive intervals Promo2 is started, naming the months the promotion is started anew. E.g. "Feb,May,Aug,Nov" means each round starts in February, May, August, November of any given year for that store  


# Exploratory data analysis
## read data and check details
```{r}
# read data
train <- read.csv("train.csv")
test <- read.csv("test.csv")
store <- read.csv("store.csv")
# check data dimensions
dim(train)
dim(test)
dim(store)
# check variable details
str(train)
str(test)
str(store)
# check the first 6 rows of the data
head(train)
head(test)
head(store)
# check the last 6 rows of the data
tail(train)
tail(test)
tail(store)
```

## merge data and transform data type
```{r}
# merge data
train <- merge(train,store,by="Store")
test <- merge(test,store,by="Store")
# transform the data type
# train
train$Store <- as.factor(train$Store)
train$DayOfWeek <- as.factor(train$DayOfWeek)
train$Open <- as.factor(train$Open)
train$Promo <- as.factor(train$Promo)
train$StateHoliday <- as.factor(train$StateHoliday)
train$SchoolHoliday <- as.factor(train$SchoolHoliday)
train$Assortment <- as.factor(train$Assortment)
train$StoreType <- as.factor(train$StoreType)
train$Promo2 <- as.factor(train$Promo2)
train$PromoInterval <- as.factor(train$PromoInterval)
train$Date <- as.Date(train$Date)

# test
test$Store <- as.factor(test$Store)
test$DayOfWeek <- as.factor(test$DayOfWeek)
test$Open <- as.factor(test$Open)
test$Promo <- as.factor(test$Promo)
test$StateHoliday <- as.factor(test$StateHoliday)
test$SchoolHoliday <- as.factor(test$SchoolHoliday)
test$Assortment <- as.factor(test$Assortment)
test$StoreType <- as.factor(test$StoreType)
test$Promo2 <- as.factor(test$Promo2)
test$PromoInterval <- as.factor(test$PromoInterval)
test$Date <- as.Date(test$Date)

# extract the month information
train$Month <- as.factor(format(train$Date, "%m"))
test$Month <- as.factor(format(test$Date, "%m"))
# extract year information
train$Year <- as.factor(format(train$Date, "%Y"))
test$Year <- as.factor(format(test$Date, "%Y"))
# extract day information
train$Day <- as.factor(format(train$Date, "%d"))
test$Day <- as.factor(format(test$Date, "%d"))

```
## save preprocessed data
```{r}
save(train, test, file = "Rossmann.RData")
```

## explore data  
### data summaries
```{r}
# summary of data
summary(train)
summary(test)
```

### Visulize data in train
```{r}
# load package
library(dplyr)
library(ggplot2)
# average sales in all stores
train %>% 
  group_by(Store) %>% summarize(AverageSales=mean(Sales)) %>%
  ggplot(aes(x=Store, y=AverageSales)) + geom_point()
# average sales for different store types
train %>% 
  group_by(StoreType, Store) %>% summarize(AverageSales=mean(Sales)) %>%
  ggplot(aes(x=Store, y=AverageSales)) + geom_point(aes(colour=StoreType))
# average sales for different assortment types
train %>% 
  group_by(Assortment, Store) %>% summarize(AverageSales=mean(Sales)) %>%
  ggplot(aes(x=Store, y=AverageSales)) + geom_point(aes(colour=Assortment))

# take store number 262 as an example to view the sales 
# different school holidays
train %>% subset(Store=="262") %>% ggplot(aes(y=Sales, x=Date)) + geom_point(aes(colour=SchoolHoliday))
# different state holidays
train %>% subset(Store=="262") %>% ggplot(aes(y=Sales, x=Date)) + geom_point(aes(colour=StateHoliday))
# different promotion status
train %>% subset(Store=="262") %>% ggplot(aes(y=Sales, x=Date)) + geom_point(aes(colour=Promo))
# different day of week
train %>% subset(Store=="262") %>% ggplot(aes(y=Sales, x=Date)) + geom_point(aes(colour=DayOfWeek))
# different month
train %>% subset(Store=="262") %>% ggplot(aes(y=Sales, x=Date)) + geom_point(aes(colour=Month))
```

